#Analysis

```{r setup, include=FALSE}
library(tidyr); library(dplyr); library(ggplot2)
learn<- read.csv("learn.csv")

```

## Introduction

In this final project I will study the data frame learning2014 introduced in the course *Introduction to Open Data Science*, or in short *IODS*. The data frame contains answers from feedback questionnare and exam points from course Introduction to Statistics held in 2014. 

```{r cars, echo=FALSE}
summary(learn)

dim(learn)
```

The data frame has 183 observations and 9 variables. Gender and age are self explanatory variables. Deep, stra and surf refer to the questions asked: deep, strategic and surface questions. Attitude contains scores from questions regarding attitude towards statistics and interest in the same way scores from questions regarding interest towards statistics. Points are the exam points gained by the student. The last variable high_points is a binomial variable containing value true, if the mean score calculated from points and attitude is higher than 26 (mean from all observations).

To study this data frame, I will introduce a hypothesis:

  <span style="color:#159957">*"Students who have on average high exam points and score high on attitude towards statistics also have high interest in statistics."*
  
This means that I will focus on variables interest and high_points in the following analysis. 

#Looking closer


```{r variables, echo=FALSE}
par(mfrow=c(2,3))

for (i in 1:ncol(learn)) {
  if(colnames(learn)[i] == "high_points"){
    high <- table(learn$high_points)
    barplot(high, main=colnames(learn)[i],  ylab = "Count",col = c("#159957"))
    break
  }else{
  plot(learn[,i], main=colnames(learn)[i],
       ylab = "Count", col = "#159957", las = 2)
  }
}

```

Looking at the variables individually there can be seen some trends in them. Clearly most of the participants are colse to aged 20 and there were more female than men. Attitude and strategic are scattered widely and cannot be further explained here. Most of the students got 15-20 points from their exam. This means that assuming passing required 15 to 18 points, most of the students who answered to the questionnare passed the exam. Interest towards statistics is grouped mostly near 8/10 as is expected (would be quite odd if most of the students of statistics are not interested at all).


```{r versus, echo=FALSE}
par(mfrow=c(2,2))

t1 <- table(high_points = learn$high_points, interest=learn$interest)
t2 <- table(high_points = learn$high_points, gender=learn$gender)
t3 <- table(high_points = learn$high_points, age=learn$age)

barplot(t1, main="High points vs Interest",  ylab = "Count",col = c("#159957", "#155D93"))
barplot(t2, main="High points vs Gender",  ylab = "Count",col = c("#159957", "#155D93"))
barplot(t3, main="High points vs Age",  ylab = "Count",col = c("#159957", "#155D93"))

```

#Logistic regression



#Odds ratio